---
title: "Data 624 HW2"
author: "V Patel"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
    highlight: tango
    theme: united
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE}
if (!require("fpp2")) install.packages("fpp2")
```

### 3.1 

For the following series, find an appropriate Box-Cox transformation in order to stabilise the variance.

* `usnetelec`
* `usgdp`
* `mcopper`
* `enplanements`

#### `usnetelec`

```{r}
autoplot(usnetelec)
box_lamda_ulec<- BoxCox.lambda(usnetelec)
autoplot(BoxCox(usnetelec, box_lamda_ulec))
```

`r box_lamda_ulec` is suggested lambda. 


#### `usgdp`

```{r}
autoplot(usgdp)
box_lamda_udp<- BoxCox.lambda(usgdp)
autoplot(BoxCox(usgdp, box_lamda_udp))
```

`r box_lamda_udp` is suggested lambda. 

#### `mcopper`

```{r}
autoplot(mcopper)
box_lamda_mcopper<- BoxCox.lambda(mcopper)
autoplot(BoxCox(mcopper, box_lamda_mcopper))
```

`r box_lamda_mcopper` is suggested lambda. 

#### `enplanements`

```{r}
autoplot(enplanements)
box_lamda_enpl<- BoxCox.lambda(enplanements)
autoplot(BoxCox(enplanements, box_lamda_enpl))
```

`r box_lamda_enpl` is suggested lambda.

### 3.2

Why is a Box-Cox transformation unhelpful for the cangas data?

```{r}
autoplot(cangas)
autoplot(BoxCox(cangas, BoxCox.lambda(cangas)))
```

I don't see an improvement on Box-Cox transformation. Hence,  Box-Cox transformation is not helpful.

### 3.3

What Box-Cox transformation would you select for your retail data (from Exercise 3 in Section 2.10)?

```{r}
library(readxl)
retaildata <- readxl::read_excel("C:/Users/patel/Documents/Data_624/retail.xlsx", skip=1)

myts <- ts(retaildata[,"A3349398A"], frequency=12, start=c(1982,4))
autoplot(myts)
```

```{r}
(lambda_retail <- BoxCox.lambda(myts))
autoplot(BoxCox(myts,lambda_retail))
```

`r lambda_retail` is suggested lambda value.

 
### 3.8

For your retail time series (from Exercise 3 in Section 2.10):

A. Split the data into two parts using

```{r}
myts.train <- window(myts, end=c(2010,12))
myts.test <- window(myts, start=2011)
```


B. Check that your data have been split appropriately by producing the following plot.

```{r}
autoplot(myts) +
  autolayer(myts.train, series="Training") +
  autolayer(myts.test, series="Test")

```



C. Calculate forecasts using snaive applied to myts.train.

```{r}
fc <- snaive(myts.train)
```



D. Compare the accuracy of your forecasts against the actual values stored in myts.test.

```{r}
accuracy(fc,myts.test)
```



E. Check the residuals.

```{r}
checkresiduals(fc)
```


Do the residuals appear to be uncorrelated and normally distributed?

**Nope, it does not correlated with each other and not normally distributed**

F. How sensitive are the accuracy measures to the training/test split?

It seems Mean Error is highly sensitive, RMSE, MAE, MPE, MASE are sensitive, and MAPE and ACF1 are not sensitive 