---
title: "Data 624 HW3"
author: "V Patel"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
    highlight: tango
    theme: united
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
if (!require("fpp2")) install.packages("fpp2")
if (!require("seasonal")) install.packages("seasonal")

```

### 6.2 The `plastics` data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.


#### A. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?

```{r fig.height=4, fig.width=6}
autoplot(plastics) + ylab("Monthly Sales")
```

**There is a increasing trend in the timeseries. All sales are at peak in summer**

#### B. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

```{r fig.height=4, fig.width=6}
fit <- decompose(plastics, type="multiplicative")

trend_indices <- fit$trend
seasonal_indices <- fit$seasonal

autoplot(trend_indices)
```

**The results in of the multiplicative decomposition show that there is an increasing trend till 5 years. After that we can see decreasing it is decresing.**

```{r fig.height=4, fig.width=6}
autoplot(seasonal_indices)
```

**seasonal_indices is identical in thoughout all years**


#### C. Do the results support the graphical interpretation from part a?

**Yes, Results does support the graphical interpretation. summer months does have sales higher than other seasonal**

#### D. Compute and plot the seasonally adjusted data.

```{r}
adjusted <- seasadj(fit)
plot(adjusted)
```

#### E. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

```{r}
obs_plast <- plastics
obs_plast[17] = obs_plast[17] + 500

fit2 <- decompose(obs_plast, type="multiplicative")

change_obes <- seasadj(fit2)
plot(change_obes)
```

**There is sudden spike in plot**

#### F. Does it make any difference if the outlier is near the end rather than in the middle of the time series?

```{r}
obs_plast1 <- plastics
obs_plast1[55] = obs_plast[55] + 500

fit3 <- decompose(obs_plast1, type="multiplicative")

change_obes_end <- seasadj(fit3)
plot(change_obes_end)
```

**It appears that it doesn't matter where the outlier occurs. we see a identical spike at the end like seen in previous plot**


### 6.3 Recall your retail time series data (from Exercise 3 in Section 2.10). Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?


```{r message=FALSE, warning=FALSE}
library(readxl)
library(seasonal)

retaildata <- readxl::read_excel("C:/Users/patel/Documents/Data_624/retail.xlsx", skip=1)

myts <- ts(retaildata[,"A3349398A"], frequency=12, start=c(1982,4))
myts %>% seas(x11="") -> myts_x11
autoplot(myts_x11) +
  ggtitle("X11 decomposition")
```

**The X11 decomposition method reveals a smooth upward trend. The remainder plot describes that after 1992 trend and seasonality appear consistent. The unusual features that I had not noticed previously was there is sudden spike at 1986 in remainder plot despite there is consistancy in tread before and after that. ** 


