---
title: "Data 624 HW7"
author: "V Patel"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    highlight: tango
    theme: united
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE}
if (!require("fpp2")) install.packages("fpp2")
if (!require("caret")) install.packages("caret")
if (!require("AppliedPredictiveModeling")) install.packages("AppliedPredictiveModeling")
```


### 6.2 Developing a model to predict permeability (see Sect 1.4) could save significant resources for a pharmaceutical company, while at the same time more rapidly identifying molecules that have a sufficient permeability to become a drug:

#### (a) Start R and use these commands to load the data:

```{r}
library(AppliedPredictiveModeling)
data(permeability)
head(permeability)
```


#### (b) The fingerprint predictors indicate the presence or absence of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure. Filter out the predictors that have low frequencies using the `nearZeroVar` function from the `caret` package. How many predictors are left for modeling?

```{r warning=FALSE}
library(caret)

dim(fingerprints)
fingerprints_nZ <- fingerprints[, -nearZeroVar(fingerprints)]
dim(fingerprints_nZ)
```

#### (c) Split the data into a training and a test set, pre-process the data, and tune a PLS model. How many latent variables are optimal and what is the corresponding resampled estimate of $R^2$ ?

```{r }
library(pls)

set.seed(17)

ctrl <- trainControl(method = "cv", number = 10)

trainingRows <- createDataPartition(permeability, 
                                    p = 0.75, 
                                    list = FALSE)

trainFingerprints <- fingerprints_nZ[trainingRows,]
trainPermeability <- permeability[trainingRows,]

testFingerprints <- fingerprints_nZ[-trainingRows,]
testPermeability <- permeability[-trainingRows,]


plsTune <- train(x = trainFingerprints, y = log10(trainPermeability),
                 method = "pls",
                 tuneGrid = expand.grid(ncomp = 1:15),
                 trControl = ctrl)

plsTune

plot(plsTune)

summary(plsTune$finalModel)
```

**PLS model is 7 is the optimal model. This captures 68.17% of the variation in the predictors and 79.33% of the variation**


#### d) Predict the response for the test set. What is the test set esitmate of $R^2$?


```{r}
pls.pred <- predict(plsTune, newdata = testFingerprints)
postResample(pred = pls.pred, obs = testPermeability)
```

The $R^2$ of the test set prediction is `0.4324274` 


#### e) Try building other models discussed in this chapter. Do any have better predictive performance?

```{r warning=FALSE}
ridgeGrid <- data.frame(lambda = seq(0.02, .35, length = 9))
ridgeTune <- train(trainFingerprints, trainPermeability, method = "ridge",
                   tuneGrid = ridgeGrid, 
                   trControl = ctrl,
                   preProc = c("center", "scale"))
ridgeTune
```



